---
title: "Capstone Project"
author: "Alex V"
date: "May 13, 2015"
output: 
  html_document: 
    keep_md: yes
---

##Import 
```{r}
library(readr)
raw <- read_csv("C:/Users/Alex/1371_traffic_volumes.csv")
raw2 <- raw[-1:-2,] #remove extraneous rows (French and blank row) to allow "readr" to reparse variables types correctly
raw3 <- type_convert(raw2)
str(raw3) #check datatypes
```

##1: Initial Cleaning
```{r echo = FALSE}
library(magrittr)
library(dplyr)

#rename columns
library(dplyr)
renamed.raw <- rename(raw3, "lhrs" = LHRS,"OS" =`O/S`, "year" = Year, "hwy.num" = `Hwy #`, "hwy.let" = `Hwy Let`, "hwy.type" = `Hwy Type (Text)`, "location" = `Location Description`, "reg" = Reg, "section.len" = `Section Length`, "conn.link.len" = `Connecting Link Length`, "sec.description" = `Secondary Description (for Connecting Links, Regional Boundarys,etc)`, "travel.pattern" = `Travel Pattern`, "dhv.percent" = `DHV %`, "directional.split" = `Directional Split`, "aadt" = AADT, "annual.aadt.change" = `Yearly % Change in AADT`, "decade.change" = `10 Year % Change in AADT`, "sadt" = SADT, "sawdt" = SAWDT, "wadt" = WADT)

#checking for NAs
length(renamed.raw$lhrs)
length(renamed.raw$year)
length(renamed.raw$aadt)
sum(is.na(renamed.raw$aadt))
sum(is.na(renamed.raw$sadt))
sum(is.na(renamed.raw$sawdt))
sum(is.na(renamed.raw$wadt))
sum(renamed.raw$year=="9999")
#there are 553 NA entries for aadt. year is also set to 9999 for these entries

#removing NAs and duplicate entries
fil.data <- renamed.raw %>%
  distinct() %>%
  filter(year!="9999")

length(fil.data$year)
length(fil.data$aadt)

glimpse(fil.data)

#convert travel.pattern and region to factor
fil.data$travel.pattern <-  factor(fil.data$travel.pattern)
fil.data$reg <- factor(fil.data$reg)
is.factor(fil.data$travel.pattern)

#grouping data by regions: Central, Eastern, Northeastern, Northwestern, West
central <- filter(fil.data, reg == "CR")
eastern <- filter(fil.data, reg == "ER")
noreast <- filter(fil.data, reg == "NE")
norwest <- filter(fil.data, reg == "NW")
west <- filter(fil.data, reg == "SW")
 
nrow(central)
nrow(eastern)
nrow(noreast)
nrow(norwest)
nrow(west)
```

There are `r sum(renamed.raw$year=="9999")` incomplete entries. These entries are set to Year "9999" and have no values for AADT and the related variables.

There are `r length(fil.data$aadt)` remaining observations

##2: Exploratory

```{r}
library(pastecs)
library(dplyr)
explor.stat <- select (fil.data, aadt:wadt)
stat.desc(explor.stat)
```

```{r}
library(psych)
describe(fil.data)
#cor(renamed.raw)
```

```{r}
#MTO divides the highways into four categories: the king's highways, the secondary highways, the tertiary roads, and selected 7000 series highways. the king's highways will be further grouped into the 400 series highways +  the QEW.

library(dplyr)

#select for the four highway types in the dataset
king <- filter(fil.data, hwy.type == "King")
secon <- filter(fil.data, hwy.type == "Sec")
fwy <- filter(fil.data, hwy.type == "Fwy")
tert <- filter(fil.data, hwy.type == "Tert")

#select for the 400, 401, 402, 403, 404, 405, 406, 407, 409, 410, 416, 417, 420, 427
target <- c("400", "401", "402", "403", "404", "405", "406", "407", "409", "410", "416", "417", "420", "427")
four <- filter(fil.data, hwy.num %in% target)

#select for the remaining non-400 series freeways (the QEW) 
qew <- filter(fwy, !(hwy.num %in% target))
```

```{r}
summary(lm(sadt~aadt, data=fil.data))
plot(fil.data$sadt~fil.data$aadt, xlab="SADT", ylab="AADT", main="Plot of SADT and AADT")
```

```{r}
fil.data %>%
  group_by(hwy.type) %>%
  ggvis (~year, ~aadt, stroke = ~factor(hwy.type)) %>%
  layer_smooths() %>%
  title= "Year ~ AADT"
```

Distribution of AADT, SADT, SAWDT, and WADT
```{r}
datadis <- select(fil.data, aadt, sadt, sawdt, wadt)
boxplot(datadis,
        main="Distribution of Data",
        ylab="Value",
        xlab="Data Type")
```

Histograms of Highway Type
```{r}
fil.data %>%
  group_by(hwy.type) %>%
  ggvis (~year, ~aadt, fill=~hwy.type) %>%
  layer_histograms() 
```

```{r}
fil.data %>%
group_by(hwy.type) %>%
ggvis(~year, ~aadt, fill = ~hwy.type) %>%
layer_densities()
```

```{r echo = FALSE}
#From R Cookbook: http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/

# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


```{r}
#describe distribution of patterns by region
#central
library(ggvis)
par(mfrow=c(3,2))
ggvis(central, ~aadt, fill:= "#fff8dc") %>%
  layer_histograms() %>%
  add_axis("x", title = "AADT") %>%
  add_axis("y", title = "Count") %>%

library(ggplot2)
  
p1 <- qplot(central$aadt,
      geom="histogram",
      main = "Histogram for Central",
      xlab = "AADT")

p2 <- qplot(eastern$aadt,
      geom="histogram",
      main = "Histogram for Eastern",
      xlab = "AADT")

p3 <- qplot(noreast$aadt,
      geom="histogram",
      main = "Histogram for Northeastern",
      xlab = "AADT")

p4 <- qplot(norwest$aadt,
      geom="histogram",
      main = "Histogram for Northwestern",
      xlab = "AADT")

p5 <- qplot(west$aadt,
      geom="histogram",
      main = "Histogram for West",
      xlab = "AADT")

p6 <- qplot(fil.data$aadt,
      geom="histogram",
      main = "Histogram for All Regions",
      xlab = "AADT")

multiplot(p1, p2, p3, p4, p5, p6, cols=3)


```


#look for areas that have large seasonal variations

##3: Visualizations

```{r}
library(ggvis)
library(ggplot2)
```

```{r}
fwy %>% 
  ggvis(~travel.pattern, ~aadt) %>%
  layer_smooths() %>%
  layer_model_predictions(model = "lm", stroke := "red", se = TRUE)

king %>% 
  ggvis(~year, ~aadt) %>%
  layer_smooths() %>%
  layer_model_predictions(model = "lm", stroke := "red", se = TRUE)

secon %>% 
  ggvis(~year, ~aadt) %>%
  layer_smooths() %>%
  layer_model_predictions(model = "lm", stroke := "red", se = TRUE)

tert %>% 
  ggvis(~year, ~aadt) %>%
  layer_smooths() %>%
  layer_model_predictions(model = "lm", stroke := "red", se = TRUE)
```

```{r}
#Distribution of Travel by Pattern
fwy %>% 
  ggvis(~travel.pattern, ~aadt) %>%
  layer_histograms()

fwyhist <- select(fwy, travel.pattern, aadt)
hist(fwyhist$aadt)


```



##4: Predictions

```{r}
library(dplyr)
library(corrplot)
cordata <- select(fil.data, aadt, sadt, sawdt, wadt)
dhvcor <- select(fil.data, dhv.percent, aadt, sadt, sawdt, wadt)
cor(cordata)
cor(dhvcor)
corrplot(cor(cordata))
corrplot(cor(dhvcor))
plot(cordata)
```

```{r}
model_ulm <- lm (aadt ~ travel.pattern, data=fil.data)
summary(model_ulm)
plot(model_ulm)

fil.data2 <- fil.data

fil.data2 <- within(fil.data2, {
    travel.patternC <- C(travel.pattern, aadt)
    print(attributes(travel.patternC))
})

```

```{r}
model_ulm <- lm (aadt ~ travel.pattern, data=fwy)
summary(model_ulm)
plot(model_ulm)

fwy %>% 
  ggvis(~travel.pattern, ~aadt) %>%
  layer_histograms() %>%
  layer_model_predictions()

ggplot(model_ulm)
```


#rank data

```{r, include=FALSE}
   # add this chunk to end of mycode.rmd
   file.rename(from="scripts/mycode.md", 
               to="README.md")
```

##GIS Mapping with ORN

```{r}
> orndbf <- read.dbf("C:/Users/Alex/ORN_SEGMENT_WITH_ADDRESS.dbf")
> str(orndbf)
```

